Optical Detection of Paramagnetic Defects in a CVD-grown Diamond

### 2021-02-26 Mes papiers/CR-CVD/Reviewer_answer.tex

- On page 2, the reference to Fig. 1-b) has a typo.

- The magnetic field axis in Figs. 3 and 4 have no units.

- The legend in Fig. 4 should read ‘13C-NV’ instead of 13C as explained in the text.
\end{quote}

\subsection*{Our answer}
Here is our answer to the points raised by Referee 1 :
\begin{enumerate}
\item We have added a short description of the three relevant defects of this study when describing Fig.1
\item We have added the concentration of NV and N in the description of the sample.
\item We have reformulated the sentence to make clear that we mention other cross-relaxation studies based on NV centers.
\item We have rewritten the paragraph to make it clear that this is indeed the same sample.
\item We have added information on the symmetry of the different defects with respect to the [100] axis to justify that we only take into account two transitions per defect in our model (except for the 13C-NV defect which is described in the supplementary material)
\item We added a discussion on the Gaussian vs Lorentzian line-shape in sec.V of the supplementary material. Given that the approximate shorter distance between our spins is more than 10 nm, the coupling strength between our spins would be less than 100 KHz. The width of the lines in fig.3 is therefore essentially due to the inhomogeneous Gaussian broadening of the ensemble of spins.
\item We added a paragraph at the end of the manuscript where we give a tentative estimate of the VH- and WAR1 concentration based on our measurement.
\item It would be very relevant to analyze a range of diamond samples grown under different conditions (methane concentration, temperature, growth rate, plasma power density etc.) with this new optical technique, however we currently only have access to a single CVD sample with a high enough concentration in NV centers to observe cross-relaxation effects. We have already found that NV concentration strongly varies with the growth temperature (see reference: A. Tallaire et al. Temperature dependent creation of nitrogen-vacancy centers in single crystal CVD diamond layers, Diam. \& Relat. Mat. 51 (2015) 55–60.) and we suspect that it may also influence the incorporation of other defects although we have not confirmed this so far. 
\end{enumerate}

@@

8. What is the quantization axis of the observed defects with respect to the [100] axis? Would the picture in Fig.4 still hold if one considers possible state mixings?

Some minor issues:

1. The legend label of 13C in Fig.4 (also Fig.1 in SI) is misleading. I would suggest label it with ‘NV- 13C’ instead of ‘13C’.

2. As mentioned before, I would suggest adding more information on the CVD sample such as 13C abundancy as well as estimation of defect concentrations, if possible.
\end{quote}
\subsection*{Our answer}
Here is our answer to the points raised by Referee 1 :
\begin{enumerate}
\item We have added a discussion of the 4th order polynomial in the description of fig.3, Referee 2 was mostly right on their assumption regarding the order of the polynomial. (although we have not really seen hiding of the spectral bumps even for higher order polynomial)
\item In the same section we explain that we tried to find a numerical solution for the overall decrease of PL with the magnetic field, but the solutions we found were not in agreement with our experimental data. (the decrease in our data seem too linear compared to the predicted quadratic profile of the numerical simulations)
\item We added a discussion regarding this point (similar to question 7 of Referee 1) at the end of the manuscript. It is in fact possible to extract an estimate of the concentrations with the area of the peaks (which is probably more relevant than the amplitude of the peaks in this case), but the width of the peak is mostly due to inhomogeneous broadening and not to interactions.
\item We added fabrication details for both the CVD and the HPHT sample in Sec.VI of the supplementary material. As for why the VH$^-$ and WAR1 defects were only found in the CVD sample, we think that although the irradiation procedure of the two samples was not perfectly identical, it would not be the reason for the observed difference in VH- and WAR1 which is mainly due to the initial starting defects in the material (HPHT vs CVD). CVD synthesis is performed in a hydrogen-rich environment and the presence of H contamination in the samples is very common. To the contrary HPHT diamonds do not generally contain significant amounts of H but the control of other type of impurities may be more problematic such N in substitutional and aggregated forms as well as metal impurities from the melted bath.  
\item As previously explained hydrogen is not normally found in HPHT-grown material due to the growth process that does not involve this gas. On the contrary, CVD growth requires a high fraction of hydrogen in the gas phase (95\% or more) which is dissociated in a micro-wave plasma. A small fraction of these reactive H species can get incorporated into the growing crystal depending on the growth conditions. This would explain why VH- (and also NVH) are commonly found in CVD-grown material but not in HPHT diamonds. As for WAR1, there is no agreement in the literature on its precise nature, however on the light of those results we believe that it is likely that it involves H impurities. 
\item We have added a discussion of the concentration of 13C-NV pairs in the last paragraph when quantifying the concentration of our species. As for the beyond fist-shell 13C we have clarified in the 13C section that any 13C further away than the first-shell would have an hyper-fine bonding of less than 15 MHz, making its contribution almost indistinguishable from the inhomogeneous broadening due to the ambiant spin bath.
\item We added on page 3 a clarification that scans in the [100] direction cannot be used to detect spins with ZFS too different from the NV centers, due to the transverse field depolarization of NV centers for high field values.
\item The quantization axes for the $S_z$ operator of NV, VH and WAR1 in fig.4 were taken to be one of the four $\langle 111 \rangle$ direction, which are all equivalent when the magnetic field is in the [100] direction. Although this is not clearly visible, fig.4 does take into account state mixing due to the transverse field, by diagonalizing the Hamiltonian with both the longitudinal and transverse contribution of the magnetic field. 
\end{enumerate}

### 2021-02-26 Mes papiers/CR-CVD/CVD_para.tex
tutu

\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{natbib}
\usepackage{physics}

@@

The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing \cite{Rondin_2014} and quantum information science  \cite{DOHERTY20131}.
One major reason is that it can be optically polarized and read-out and features long coherence time even at ambient conditions. 
%The NV property can also then be employed to manipulate and to detect other dark spins that are interacting with it through dipole-dipole coupling, either inside or outside the diamond. 
In order to optimize the capabilities of the NV center, its magnetic environment must be very well controlled. 
Synthetic growth of diamond crystals is now reaching a level of maturity that makes the most pristine diamond crystals almost flawless, thus eliminating the source of magnetic noise from other nearby impurities than the  NV sensor. The standard growth technique works through Chemical Vapor Deposition (CVD) of carbon atoms from a methane gas in an ultra-high vacuum environment. It offers the possibility to use crystals with 99.9\% purity, as well as using $^{12}$C enriched methane, hence removing spin fluctuations from $^{13}$C atoms and enhancing further the sensing capabilities of NV centers \cite{Achard}.

@@

%In CVD grown samples, the typical paramagnetic impurities that have been detected by EPR are hydrogen complex such as the NVH, VH, VH2, where the hydrogen comes from the methane in the reactor. Fig.  \ref{fig1} shows a schematics of the various paramagnetic defects in a CVD grown diamonds that are coupled to the Nitrogen Vacancy center in our study.
Previous cross-relaxation studies with NV$^-$ centers have been mostly carried out using heavily doped crystals grown by the High-Pressure-High-Temperature (HPHT) process \citep{Epstein,Wang, armstrong_nvnv_2010, Hall, Wickenbrock, Alfasi}. 
Recent efforts in the doping processes have also made it possible to reach NV centers concentrations in the 3 to 5 ppm range in CVD grown samples \cite{Edmonds, TALLAIRE2020421, MINDARAVA2020182}, opening a path towards detecting other paramagnetic defects in CVD grown diamonds. 
%these hydrogen-related complexes using cross-relaxation with NV centers.%Several impurities in CVD grown diamonds have been detected using Electron-Paramagnetic Resonance.  In order to purify diamond crystal and optimize their conduction and properties for prolonged qubit lifetimes. 
%Extensive work has been done in the past twenty years to detect and characterize the various  that appear in EPR. 
The sample we use in this study  correspond to the pink irradiated sample in \cite{TALLAIRE2020421}. It was grown using CVD with the addition of 500ppm of N$_2$0 to the H2/CH4 (96/4) gas phase.
Then high energy (10MeV) electron irradiation at a fluence of $2\times18$ cm$^{-2}$ and at a temperature of 900 degrees celsius was realised, giving a final concentration of NV$^-$ of about 4.6~ppm for an initial concentration of nitrogen in the sample of about 25~ppm.
It was shown that in this sample the $T_2^*$ was not degraded after annealing, yet the NV density was large enough to enable concentration dependent longitudinal relaxation \cite{TALLAIRE2020421, jarmola_longitudinal_2015}.  
%It is well-known that when the NV concentration increases above 1 ppm, the longitudinal decay of NV centers is degraded. 
%Furthermore, it has been observed that cross-relaxation amongst closely pack NVs of different classes can be observed.
%It was postulated that such a degradation of the $T_1$ time came from the coupling of NV centers to other NV-related defects.

In this study, we use a homebuilt confocal microscope that comprises a 1mW green laser, an objective with a numerical aperture of 0.25 to focus the laser onto the sample as well as to collect the NV photoluminescence (PL). The PL was filtered from the green laser using a dichroic mirror and a notch filter. It was then coupled to a multimode fiber and detected by an avalanche photodiode. The magnetic field scans were realized using a C-shaped electromagnet driven by a current generator (HP 33120A).  We then monitor the NV PL synchronously with the changes in the magnetic field.
\begin{figure}[ht]
\centering
{\includegraphics[width=\linewidth]{Scans_fig4}}
\caption{Optical detection of cross-relaxations. {Top}: NV$^-$ photoluminescence counts as a function of the magnetic field amplitude along the [100] crystalline direction (plain line) and fourth-order polynomial fit (dashed line). {Bottom} Curve obtained by subtracting the polynomial fit to the above signal (circles). The simulated cross-relaxations positions are shown by the green dashed, vertical lines. Gaussian fits to the second and third dips are shown by plain orange lines.}
\label{scan}
\end{figure}
Contrary to the more commonly employed $[ 111 ]$ direction, we scan the magnetic field along the diamond $[100]$ crystalline direction. 
Looking at the diamond structure (bottom right in Fig.  \ref{fig1}), it can be noticed that, in this direction, the four possible projections of the NV centers (or any other $C_{3v}$ defect) along the magnetic field are identical. This means that a B field scan along this direction cannot make the transition energies of the different classes of NV centers cross. Under arbitrary angles, this could otherwise result in several low field PL features \cite{van_oort_optically_1991, akhmedzhanov_magnetometry_2019} that could mask cross-relaxation from other species and also reduce the CR contrast (see section 4 of the supplementary materials for data showing scans along the [111] direction). 
The downside of this choice of the [100] direction is that at large amplitudes, the transverse component of the magnetic field depolarizes all classes of NVs, thus limiting the magnetic field range that can be employed and therefore limiting this detection process to paramagnetic defects with zero-field splittings close to the NV center's one.

@@

In order to let the three salient features detach better from the spectrum, we fitted a 4th-order polynomial to the PL evolution, without the spectral bumps, and subtract it from the data. We chose to use a 4th-order polynomial since numerical solving of the transverse-field dependent spin depolarization could not properly fit our results. The order of the polynomial fit was chosen to be 4 since higher order did not yield significantly better envelope fitting, and lower order were noticeably worse. Thanks to the subtraction, the width and contrast of each peak can now be compared and their position can be determined with greater accuracy. The second and third peaks were better fitted with Gaussian line shapes than Lorentzian ones, which is coherent with the fact that the inhomogeneous broadening of the various spin species dominate the coupling strength between the spins.

In order to attribute the three features to their respective defects, we run a similar scan (Section 1 in the Supplementary Materials) on a type Ib electron irradiated HPHT diamond crystal with a NV concentration expected to be in the 5-20 ppm range, in the same $[100]$ direction. Only the first feature appeared. 
This observation guided us to search for CVD-related defects as good candidates for the last two features. Paramagnetic defects in CVD-grown diamonds have been extensively studied using EPR  \citep{Ashfold}. These studies demonstrate that hydrogen-related complex such as the NVH, VH, VH2 complexes can all be stable in diamond. Although their composition is not always known, several zero-field splittings $D$ can be found in the literature.
Given the small difference between the NV transitions and the observed features, we only concentrate on the reported defects with ZFS that are close to the NV$^-$.
The middle column of table \ref{table}, shows the zero-field splitting for two such spin-1 defects found in \citep{cruddace2007magnetic}, namely the negatively charged hydrogen-vacancy (VH$^-$) and WAR1 defects. The latter was analyzed in EPR but its exact structure is unknown \citep{cruddace2007magnetic}.
\begin{table}[htbp]
\centering
\caption{\bf Zero-field splitting parameter $D$ for the spin-1 species in our sample}
\begin{tabular}{ccc}
\hline
$D_z$ estimation (MHz) & Cruddace's work\citep{cruddace2007magnetic} & Our work \\
\hline
NV$^-$ & 2872(7) & * \\
VH$^-$ & 2706(30) & 2694(5)  \\
WAR1 & 2466(60) & 2470(10) \\
\hline
\end{tabular}
  \label{table}
\end{table}
Fig. \ref{energy-levels} shows the frequencies of the NV$^-$, the VH$^-$ and WAR1 spin transitions as a function of the magnetic field amplitude along the $[100]$ direction. Since the NV$^-$ and VH$^-$ are $C_{3v}$ defects, and WAR1 is a pseudo-$C_{3v}$ defect \citep{cruddace2007magnetic}, all possible orientation are degenerate for this magnetic field orientation, giving only two lines corresponding to the $\ket{0} \to \ket{+1}$ and $\ket{0} \to \ket{-1}$ transitions. The points were the NV levels cross the other defects can give rise to cross-relaxation.

@@

\begin{figure}[htbp]
\centering
{\includegraphics[width=\linewidth]{Theorie_CR}}
\caption{Simulated transition energies of the various considered spins as a function of a magnetic field aligned to the [100] axis. The NV centers electronic spin transitions are shown by the plain thick line, the VH$^-$ using dashed lines, the WAR1 using dash-dotted lines and $^{13}$C-NV pair using dotted lines. The amplitudes of the magnetic field where the energy level of the NV center crosses the one of another species are represented by vertical dashed lines.}
\label{energy-levels}
\end{figure}
The remaining broader feature in the spectrum could be coming from dipolar coupling between the spin of NV centers and paramagnetic defects that are present also in HPHT-grown diamonds, such as the substitutional nitrogen, also called $P_1$ centers ($[P_1]\approx 5-20$ ppm in our sample) or the $^{13}$C atoms (natural isotopic abundance [$^{13}$C]= 1\%). 
Using the results from \cite{simanovskaia_sidebands_2013}, we first intended to correlate its spectral position to the transitions of dipolar-coupled NV and $P_1$ centers. 
 $P_1$ centers have a zero-field splitting of only 100 MHz. In order to cross one of the two NV spin transitions at a magnetic field of 20 G, they would then have to be coupled off-resonantly to the spin of nearby NV centers. The resulting pair could then be coupled resonantly to a nearby polarized NV.
The obtained eigenfrequencies of the dipolar coupled P$_1$-NV pair that we extracted were all inconsistent with the spectral positions of our observed peak within our spectral resolution error ($\approx$1G) (see Section 3 of the supplementary materials). This leads us to consider instead the nuclear spin of the $^{13}$C atoms as the most likely candidate.

The nuclear spin of the $^{13}$C atoms do not have a ZFS and its gyromagnetic factor is four orders of magnitude lower than that of the electron. Here again, in order to give a CR contribution to the spectrum at modest magnetic fields, the nuclear spin of the $^{13}$C would first have to be strongly coupled to an NV center. The resulting pair could then be co-resonant with the spin transitions of a bare NV center, as depicted in Fig. 1. This situation can manifest itself when the nuclear spin of $^{13}$C atoms are a few shells distant from the nitrogen-vacancy centers. A strong hyperfine coupling rate of 130 MHz can for instance be reached with a $^{13}$C that is only one shell away from the NV center. In contrast, higher order shells have a maximum hyperfine coupling of 15 MHz \cite{smeltzer201113c} making them hard to distinguish from the inhomogeneous broadening.
The modeling of this interaction is presented in the section 2 of the supplementary materials. The dotted lines in Fig. \ref{energy-levels} show the four transition frequencies of the first shell $^{13}$C coupled to an NV center. The crossings between the $|m_s=0\rangle$ to $|m_s=\pm 1\rangle$ NV transition and the four transitions of this  $^{13}$C-NV pair occur at magnetic field values around 20~G. 
We added these four lines to the experimental spectrum shown in Fig. \ref{scan}. A good agreement is found between the magnetic field at which of these four transitions cross the NV and the broad 20~G feature, letting us conclude that it is the result of a CR between an unpolarized $^{13}$C-NV pair and a polarized NV center.

Because the sample we use has a natural 1.1\% abundance of $^{13}$C, we can actually use the first peak as a way to calibrate the concentration of VH$^-$ and WAR1 in our sample. Because there are three possible sites for a first-shell $^{13}$C-NV pair, we can estimate that the concentration of $^{13}$C-NV pair is about 3.3\% of the total concentration of NV centers, i.e. $\sim$135~ppb. We can then compare the area of the second and third peak to the first one to have a first order estimation of the concentration of VH$^-$ and WAR1. We find [VH$^-$]$\approx$195~ppb and [WAR1]$\approx$124~ppb. We decided to compare the area of the peaks instead of the amplitudes since the width of the peak is mostly due to inhomogeneous spread of the spins, and not to the spin-spin interaction strength. It should be noted that a severe drawback of this method is the loss of contrast for higher magnetic field amplitude, meaning in this case that the WAR1 concentration is probably undervalued.

>>>>

\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{natbib}
 \usepackage{soulutf8}
\usepackage{physics}

\usepackage{xcolor, soul}
\sethlcolor{green}
 \sethlcolor{green}
    \newcommand{\hlc}[2][yellow]{{\sethlcolor{#1} \hl{#2}}}

@@

%Document
%Voici un \hl{texte surligné en vert}
The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing \cite{Rondin_2014} and quantum information science  \cite{DOHERTY20131}.
One major reason is that it can be optically polarized and read-out and features long coherence time even at ambient conditions. 
%The NV property can also then be employed to manipulate and to detect other dark spins that are interacting with it through dipole-dipole coupling, either inside or outside the diamond. 
In order to optimize the capabilities of the NV center, its magnetic environment must be very well controlled. 
Synthetic growth of diamond crystals is now reaching a level of maturity that makes the most pristine diamond crystals almost flawless, thus eliminating the source of magnetic noise from other nearby impurities than the  NV sensor. The standard growth technique works through Chemical Vapor Deposition (CVD) of carbon atoms from a methane gas in an ultra-high vacuum environment. It offers the possibility to use crystals with 99.9\% purity, as well as using $^{12}$C enriched methane, hence removing spin fluctuations from $^{13}$C atoms and enhancing further the sensing capabilities of NV centers \cite{Achard}.

@@

%In CVD grown samples, the typical paramagnetic impurities that have been detected by EPR are hydrogen complex such as the NVH, VH, VH2, where the hydrogen comes from the methane in the reactor. Fig.  \ref{fig1} shows a schematics of the various paramagnetic defects in a CVD grown diamonds that are coupled to the Nitrogen Vacancy center in our study.
Previous cross-relaxation studies with NV$^-$ centers have been mostly carried out using heavily doped crystals grown by the High-Pressure-High-Temperature (HPHT) process \citep{Epstein,Wang, armstrong_nvnv_2010, Hall, Wickenbrock, Alfasi}. 
 Recent efforts in the doping processes have also made it possible to reach NV centers concentrations in the 3 to 5 ppm range in CVD grown samples \cite{Edmonds, TALLAIRE2020421, MINDARAVA2020182}, opening a path towards detecting other paramagnetic defects in CVD grown diamonds. 
%these hydrogen-related complexes using cross-relaxation with NV centers.%Several impurities in CVD grown diamonds have been detected using Electron-Paramagnetic Resonance.  In order to purify diamond crystal and optimize their conduction and properties for prolonged qubit lifetimes. 
%Extensive work has been done in the past twenty years to detect and characterize the various  that appear in EPR. 
The sample we use in this study  correspond to the pink irradiated sample in \cite{TALLAIRE2020421}. It was grown using CVD with the addition of 500ppm of N$_2$0 to the H2/CH4 (96/4) gas phase.
Then high energy (10MeV) electron irradiation at a fluence of $2\times18$ cm$^{-2}$ and at a temperature of 900 degrees celsius was realized,  giving a final concentration of NV$^-$ of about 4.6~ppm for an initial concentration of nitrogen in the sample of about 25~ppm.
It was shown that in this sample the $T_2^*$ was not degraded after annealing, yet the NV density was large enough to enable concentration dependent longitudinal relaxation \cite{TALLAIRE2020421, jarmola_longitudinal_2015}.  
%It is well-known that when the NV concentration increases above 1 ppm, the longitudinal decay of NV centers is degraded. 
%Furthermore, it has been observed that cross-relaxation amongst closely pack NVs of different classes can be observed.
%It was postulated that such a degradation of the $T_1$ time came from the coupling of NV centers to other NV-related defects.

In this study, we use a homebuilt confocal microscope that comprises a 1mW green laser, an objective with a numerical aperture of 0.25 to focus the laser onto the sample as well as to collect the NV photoluminescence (PL). The PL was filtered from the green laser using a dichroic mirror and a notch filter. It was then coupled to a multimode fiber and detected by an avalanche photodiode. The magnetic field scans were realized using a C-shaped electromagnet driven by a current generator (HP 33120A).  We then monitor the NV PL synchronously with the changes in the magnetic field.
\begin{figure}[ht]
\centering
{\includegraphics[width=\linewidth]{Scans_fig4}}
\caption{Optical detection of cross-relaxations. {Top}: NV$^-$ photoluminescence counts as a function of the magnetic field amplitude along the [100] crystalline direction (plain line) and fourth-order polynomial fit (dashed line). {Bottom} Curve obtained by subtracting the polynomial fit to the above signal (circles). The simulated cross-relaxations positions are shown by the green dashed, vertical lines. Gaussian fits to the second and third dips are shown by plain orange lines.}
\label{scan}
\end{figure}
Contrary to the more commonly employed $[ 111 ]$ direction, we scan the magnetic field along the diamond $[100]$ crystalline direction. 
Looking at the diamond structure (bottom right in Fig.  \ref{fig1}), it can be noticed that, in this direction, the four possible projections of the NV centers (or any other $C_{3v}$ defect) along the magnetic field are identical. This means that a B field scan along this direction cannot make the transition energies of the different classes of NV centers cross. Under arbitrary angles, this could otherwise result in several low field PL features \cite{van_oort_optically_1991, akhmedzhanov_magnetometry_2019} that could mask cross-relaxation from other species and also reduce the CR contrast (see section 4 of the Supplementary Materials for data showing scans along the [111] direction). 
The downside of this choice of the [100] direction is that at large amplitudes, the transverse component of the magnetic field depolarizes all classes of NVs, thus limiting the magnetic field range that can be employed and restricting this detection method to paramagnetic defects with zero-field splittings close to the NV center's.

@@

To let the three salient features detach better from the spectrum, we fitted a 4th-order polynomial to the PL evolution, without the spectral bumps, and subtract it from the data. Further details on the fitting choices are provided in Sec.V of the Supplementary material.

In order to attribute the three features to their respective defects, we run a similar scan (Section 1 in the Supplementary Materials) on a type Ib electron irradiated HPHT diamond crystal with a NV concentration expected to be in the 5-20 ppm range, in the same $[100]$ direction. Only the first feature appeared. 
This observation guided us to search for CVD-related defects as good candidates for the last two features. Paramagnetic defects in CVD-grown diamonds have been extensively studied using EPR  \citep{Ashfold}. These studies demonstrate that hydrogen-related complex such as the NVH, VH, VH2 complexes can all be stable in diamond. Although their composition is not always known, several zero-field splittings $D$ can be found in the literature.
Given the small difference between the NV transitions and the observed features, we only concentrate on the reported defects with ZFS that are close to the NV$^-$.
The middle column of table \ref{table}, shows the zero-field splitting for two such spin-1 defects found in \citep{cruddace2007magnetic}, namely the negatively charged hydrogen-vacancy (VH$^-$) and WAR1 defects. The latter was analyzed in EPR but its exact structure is unknown \citep{cruddace2007magnetic}.
\begin{table}[htbp]
\centering
\caption{\bf Zero-field splitting parameter $D$ for the spin-1 species in our sample}
\begin{tabular}{ccc}
\hline
$D_z$ estimation (MHz) & Cruddace's work\citep{cruddace2007magnetic} & Our work \\
\hline
NV$^-$ & 2872(7) & * \\
VH$^-$ & 2706(30) & 2694(5)  \\
WAR1 & 2466(60) & 2470(10) \\
\hline
\end{tabular}
  \label{table}
\end{table}
Fig. \ref{energy-levels} shows the frequencies of the NV$^-$, the VH$^-$ and WAR1 spin transitions as a function of the magnetic field amplitude along the $[100]$ direction. Since the NV$^-$ and VH$^-$ are $C_{3v}$ defects, and the WAR1 is a pseudo-$C_{3v}$ defect \citep{cruddace2007magnetic}, all possible orientation are degenerate for this magnetic field orientation, giving only two lines corresponding to the $\ket{0} \to \ket{+1}$ and $\ket{0} \to \ket{-1}$ transitions. The points were the NV levels cross the other defects can give rise to cross-relaxation.

@@

\begin{figure}[htbp]
\centering
{\includegraphics[width=\linewidth]{Theorie_CR}}
\caption{Simulated transition energies of the various considered spins as a function of a magnetic field aligned to the [100] axis. The NV centers electronic spin transitions are shown by the plain thick line, the VH$^-$ using dashed lines, the WAR1 using dash-dotted lines and $^{13}$C-NV pair using dotted lines. The amplitudes of the magnetic field where the energy level of the NV center crosses the one of another species are represented by vertical dashed lines.}
\label{energy-levels}
\end{figure}
The remaining broader feature in the spectrum could be coming from dipolar coupling between the spin of NV centers and paramagnetic defects that are present also in HPHT-grown diamonds, such as the substitutional nitrogen, also called $P_1$ centers ($[P_1]\approx 5-20$ ppm in our sample) or the $^{13}$C atoms (natural isotopic abundance [$^{13}$C]= 1\%). 
Using the results from \cite{simanovskaia_sidebands_2013}, we first intended to correlate its spectral position to the transitions of dipolar-coupled NV and $P_1$ centers. 
 $P_1$ centers have a zero-field splitting of only 100 MHz. In order to cross one of the two NV spin transitions at a magnetic field of 20 G, they would then have to be coupled off-resonantly to the spin of nearby NV centers. The resulting pair could then be coupled resonantly to a nearby polarized NV.
The obtained eigenfrequencies of the dipolar coupled P$_1$-NV pair that we extracted were all inconsistent with the spectral positions of our observed peak within our spectral resolution error ($\approx$1G) (see Section 3 of the Supplementary Materials). This leads us to consider instead the nuclear spin of the $^{13}$C atoms as the most likely candidate.

The nuclear spin of the $^{13}$C atoms do not have a ZFS and its gyromagnetic factor is four orders of magnitude lower than that of the electron. Here again, in order to give a CR contribution to the spectrum at modest magnetic fields, the nuclear spin of the $^{13}$C would first have to be strongly coupled to an NV center. The resulting pair could then be co-resonant with the spin transitions of a bare NV center, as depicted in Fig. 1. This situation can manifest itself when the nuclear spin of $^{13}$C atoms are a few shells distant from the nitrogen-vacancy centers. A strong hyperfine coupling rate of 130 MHz can for instance be reached with a $^{13}$C that is only one shell away from the NV center.  
In contrast, higher order shells have a maximum hyperfine coupling of 15 MHz \cite{smeltzer201113c} making them hard to distinguish from the inhomogeneous broadening.
The modeling of this interaction is presented in the section 2 of the Supplementary Materials. The dotted lines in Fig. \ref{energy-levels} show the four transition frequencies of the first shell $^{13}$C coupled to an NV center. The crossings between the $|m_s=0\rangle$ to $|m_s=\pm 1\rangle$ NV transition and the four transitions of this  $^{13}$C-NV pair occur at magnetic field values around 20~G. 
We added these four lines to the experimental spectrum shown in Fig. \ref{scan}. A good agreement is found between the magnetic field at which of these four transitions cross the NV and the broad 20~G feature, letting us conclude that it is the result of a CR between an unpolarized $^{13}$C-NV pair and a polarized NV center.

Because the sample we use has a very well determined concentration of $^{13}$C (natural 1.1\% abundance), we can use the first CR peak to calibrate the concentration of VH$^-$ and WAR1 in our sample. There are three possible sites for a first-shell $^{13}$C-NV pair, so the concentration of $^{13}$C-NV pair is about 3.3\% that of the NV centers, i.e. $\sim$135~ppb. Comparing the area of the second and third peak to the first one, we find [VH$^-$]$\approx$195~ppb and [WAR1]$\approx$124~ppb. Using areas is here more relevant than using amplitudes because inhomogeneous spread dominates over the spin-spin interaction. Note that the  WAR1 concentration is probably undervalued due to state mixing in the NV centers at the NV-WAR1 CR.
